% !TEX program = XeLaTeX
% !TEX encoding = UTF-8
\documentclass[UTF8,nofonts]{ctexart}
\setCJKmainfont[BoldFont=STHeiti,ItalicFont=STKaiti]{STSong}
\setCJKsansfont[BoldFont=STHeiti]{STXihei}
\setCJKmonofont{STFangsong}

\begin{document}

\chapter{工具库 Tools}

\epigraph{
Dennis Yurichev已经把这本书免费和开源 (libre)，作为对免费知识和教育的贡献。
同样，因为免费和开源的原则，我们需要免费的 (libre) 逆向工程工具来替代本书中提到
商业的专有工具。}{Richard M. Stallman}

\section{二进制文件分析 Binary analysis}

不需要运行程序的静态分析工具。

\myindex{Hiew}
\myindex{UNIX!strings}
\myindex{UNIX!xxd}
\myindex{UNIX!od}

\begin{itemize}
\item
(免费, 开源) \IT{ent}\footnote{\url{http://www.fourmilab.ch/random/}}: 熵 (entropy) 分析工具。
更多关于熵: \myref{entropy}.

\item
\label{Hiew}
\IT{Hiew}\footnote{\href{http://go.yurichev.com/17035}{hiew.ru}}:
针对于对二进制文件的小改动。

\item (免费, 开源) \IT{xxd} and \IT{od}: UNIX环境下用来显示二进制文件的标准工具。

\item (免费, 开源) \IT{strings}: *NIX环境下用来在二进制文件中寻找ASCII字符串的工具。
Sysinternals提供另一个版本\footnote{\url{https://technet.microsoft.com/en-us/sysinternals/strings}}
支持宽字符串 (UTF-16, Windows中广泛使用)。

\item (免费, 开源) \IT{Binwalk}\footnote{\url{http://binwalk.org/}}: 分析固件镜像的工具。

\item
\myindex{binary grep}
(免费, 开源) \IT{binary grep}:
用来在众多文件中寻找任意字节序列的小工具，
包括非执行文件: \BGREPURL.
\end{itemize}

\subsection{反汇编器 Disassemblers}

\myindex{IDA}
\myindex{Binary Ninja}
\myindex{BinNavi}
\myindex{objdump}

\begin{itemize}
\item \IT{IDA}. 一个老的但是免费的版本可以在这里下载
\footnote{\href{http://go.yurichev.com/17031}{hex-rays.com/products/ida/support/download\_freeware.shtml}}.
\ShortHotKeyCheatsheet: \myref{sec:IDA_cheatsheet}

\item \IT{Binary Ninja}\footnote{\url{http://binary.ninja/}}

\item (免费, 开源) \IT{zynamics BinNavi}\footnote{\url{https://www.zynamics.com/binnavi.html}}

\item (免费, 开源) \IT{objdump}: 显示和反汇编二进制文件的小命令行工具。

\item (Free, open-source) \IT{readelf}\footnote{\url{https://sourceware.org/binutils/docs/binutils/readelf.html}}:
dump information about ELF file.
\end{itemize}

\subsection{Decompilers}

There is only one known, publicly available, high-quality decompiler to C code: \IT{Hex-Rays}:\\
\href{http://go.yurichev.com/17033}{hex-rays.com/products/decompiler/}

\subsection{Patch comparison/diffing}

You may want to use it when you compare original version of some executable and patched one, in order to find
what has been patched and why.

\begin{itemize}
\item (Free) \IT{zynamics BinDiff}\footnote{\url{https://www.zynamics.com/software.html}}

\item (Free, open-source) \IT{Diaphora}\footnote{\url{https://github.com/joxeankoret/diaphora}}
\end{itemize}

\section{Live analysis}

Tools you use on a live system or during running of a process.

\subsection{Debuggers}

\myindex{\olly}
\myindex{Radare}
\myindex{GDB}
\myindex{tracer}
\myindex{LLDB}
\myindex{WinDbg}

\begin{itemize}
\item (Free) \IT{OllyDbg}.
Very popular user-mode win32 debugger\footnote{\href{http://go.yurichev.com/17032}{ollydbg.de}}.
\ShortHotKeyCheatsheet: \myref{sec:Olly_cheatsheet}

\item (Free, open-source) \IT{GDB}.
Not quite popular debugger among reverse engineers, because it's intended mostly for programmers.
Some commands: \myref{sec:GDB_cheatsheet}.
There is a visual interface for GDB, ``GDB dashboard''\footnote{\url{https://github.com/cyrus-and/gdb-dashboard}}.

\item (Free, open-source) \IT{LLDB}\footnote{\url{http://lldb.llvm.org/}}.

\item \IT{WinDbg}\footnote{\url{https://developer.microsoft.com/en-us/windows/hardware/windows-driver-kit}}:
kernel debugger for Windows.

\item (Free, open-source) \IT{Radare} \ac{AKA} rada.re \ac{AKA} r2\footnote{\url{http://rada.re/r/}}.
A GUI also exists: \IT{ragui}\footnote{\url{http://radare.org/ragui/}}.

\item (Free, open-source) \IT{tracer}.
\label{tracer}
The author often uses \IT{tracer}
\footnote{\href{http://go.yurichev.com/17338}{yurichev.com}}
instead of a debugger.

The author of these lines stopped using a debugger eventually, since all he needs from it is to spot function arguments while
executing, or registers state at some point.
Loading a debugger each time is too much, so a small utility called \IT{tracer} was born.
It works from command line, allows intercepting function execution,
setting breakpoints at arbitrary places, reading and changing registers state, etc.

N.B.: the \IT{tracer} isn't evolving, because it was developed as a demonstration tool for this book, not as everyday tool.
\end{itemize}

\subsection{Library calls tracing}

\IT{ltrace}\footnote{\url{http://www.ltrace.org/}}.

\subsection{System calls tracing}

\label{strace}
\myindex{strace}
\myindex{dtruss}
\subsubsection{strace / dtruss}

\myindex{syscall}
It shows which system calls (syscalls(\myref{syscalls})) are called by a process right now.

For example:

\begin{lstlisting}
# strace df -h

...

access("/etc/ld.so.nohwcap", F_OK)      = -1 ENOENT (No such file or directory)
open("/lib/i386-linux-gnu/libc.so.6", O_RDONLY|O_CLOEXEC) = 3
read(3, "\177ELF\1\1\1\0\0\0\0\0\0\0\0\0\3\0\3\0\1\0\0\0\220\232\1\0004\0\0\0"..., 512) = 512
fstat64(3, {st_mode=S_IFREG|0755, st_size=1770984, ...}) = 0
mmap2(NULL, 1780508, PROT_READ|PROT_EXEC, MAP_PRIVATE|MAP_DENYWRITE, 3, 0) = 0xb75b3000
\end{lstlisting}

\myindex{\MacOSX}
\MacOSX has dtruss for doing the same.

\myindex{Cygwin}
Cygwin also has strace, but as far as it's known, it works only for .exe-files
compiled for the cygwin environment itself.

\subsection{Network sniffing}

\IT{Sniffing} is intercepting some information you may be interested in.

(Free, open-source) \IT{Wireshark}\footnote{\url{https://www.wireshark.org/}} for network sniffing.
It has also capability for USB sniffing\footnote{\url{https://wiki.wireshark.org/CaptureSetup/USB}}.

Wireshark has a younger (or older) brother \IT{tcpdump}\footnote{\url{http://www.tcpdump.org/}}, simpler command-line tool.

\subsection{Sysinternals}

\myindex{Sysinternals}
(Free) Sysinternals (developed by Mark Russinovich)
\footnote{\url{https://technet.microsoft.com/en-us/sysinternals/bb842062}}.
At least these tools are important and worth studying: Process Explorer, Handle, VMMap, TCPView, Process Monitor.

\subsection{Valgrind}

(Free, open-source) a powerful tool for detecting memory leaks: \url{http://valgrind.org/}.
Due to its powerful \ac{JIT} mechanism, Valgrind is used as a framework for other tools.

% TODO network fuzzing

\subsection{Emulators}

\begin{itemize}
\item (Free, open-source) \IT{QEMU}\footnote{\url{http://qemu.org}}: emulator for various CPUs and architectures.

\item (Free, open-source) \IT{DosBox}\footnote{\url{https://www.dosbox.com/}}: MS-DOS emulator, mostly used for retrogaming.

\item (Free, open-source) \IT{SimH}\footnote{\url{http://simh.trailing-edge.com/}}: emulator of ancient computers, mainframes, etc.
\end{itemize}

\section{Other tools}

\IT{Microsoft Visual Studio Express}
\footnote{\href{http://go.yurichev.com/17034}{visualstudio.com/en-US/products/visual-studio-express-vs}}:
Stripped-down free version of Visual Studio, convenient for simple experiments.

Some useful options: \myref{sec:MSVC_options}.

\section{Something missing here?}

If you know a great tool not listed here, please drop me a note:\\
\TT{\EMAIL}.

\end{document}

