\subsubsection{Now let's get back to MSVC}

\myindex{\Cpp!exceptions}

Supposedly, Microsoft programmers needed exceptions in C, but not in \Cpp, so they added a non-standard C extension
to MSVC\footnote{\href{http://go.yurichev.com/17057}{MSDN}}.
It is not related to C++ \ac{PL} exceptions.

% FIXME russian listing:
\begin{lstlisting}[style=customc]
__try
{
    ...
}
__except(filter code)
{
    handler code
}
\end{lstlisting}

\q{Finally} block may be instead of handler code:

\begin{lstlisting}[style=customc]
__try
{
    ...
}
__finally
{
    ...
}
\end{lstlisting}


The filter code is an expression, telling whether this handler code corresponds to the exception raised.

If your code is too big and cannot fit into one expression, a separate filter function can be defined.\\
\\
There are a lot of such constructs in the Windows kernel.
Here are a couple of examples from there (\ac{WRK}):

\lstinputlisting[caption=WRK-v1.2/base/ntos/ob/obwait.c,style=customc]{OS/SEH/2/wrk_ex1.c}

\lstinputlisting[caption=WRK-v1.2/base/ntos/cache/cachesub.c,style=customc]{OS/SEH/2/wrk_ex2.c}

Here is also a filter code example:

\lstinputlisting[caption=WRK-v1.2/base/ntos/cache/copysup.c,style=customc]{OS/SEH/2/wrk_ex3.c}

Internally, SEH is an extension of the OS-supported exceptions.
But the handler function is \TT{\_except\_handler3} (for SEH3) or \TT{\_except\_handler4} (for SEH4).

The code of this handler is MSVC-related, it is located in its libraries, or in msvcr*.dll.
It is very important to know that SEH is a MSVC thing.

Other win32-compilers may offer something completely different.

\myparagraph{SEH3}

SEH3 has \TT{\_except\_handler3} 
as a handler function, and extends the \TT{\_EXCEPTION\_REGISTRATION} table, adding
a pointer to the \IT{scope table} and \IT{previous try level} variable.
SEH4 extends the \IT{scope table} 
by 4 values for buffer overflow protection.\\
\\
The \IT{scope table} is a table that consists of pointers to the filter and handler code blocks, for each nested level of \IT{try/except}.

\input{OS/SEH/2/tikz}

Again, it is very important to understand that the \ac{OS} takes care only of the \IT{prev/handle} fields, and nothing more.\\
It is the job of the \TT{\_except\_handler3} function to read the other fields and \IT{scope table}, and decide
which handler to execute and when.\\
\\
\myindex{Wine}
\myindex{ReactOS}
The source code of the \TT{\_except\_handler3} function is closed.

However, Sanos OS, which has a win32 compatibility layer, has the same
functions reimplemented, which are somewhat equivalent to those in Windows
\footnote{\url{http://go.yurichev.com/17058}}.
Another reimplementation is present in 
Wine\footnote{\href{http://go.yurichev.com/17059}{GitHub}}
and ReactOS\footnote{\url{http://go.yurichev.com/17060}}.\\
\\
If the \IT{filter} pointer is NULL, the \IT{handler} 
pointer is the pointer to the \IT{finally} code block.\\
\\
During execution, the \IT{previous try level} value in the stack changes, so
\TT{\_except\_handler3} can get information about the current level of nestedness, 
in order to know which \IT{scope table} entry to use.

\myparagraph{SEH3: one try/except block example}

\lstinputlisting[style=customc]{OS/SEH/2/2.c}

\lstinputlisting[caption=MSVC 2003,style=customasmx86]{OS/SEH/2/2_SEH3.asm}

Here we see how the SEH frame is constructed in the stack.
The \IT{scope table} is located in the \TT{CONST} segment---indeed, these fields are not to be changed.
An interesting thing is how the \IT{previous try level} variable has changed.
The initial value is \TT{0xFFFFFFFF} ($-1$).
The moment when the body of the \TT{try} 
statement is opened is marked with an instruction that writes 0 to the variable.
The moment when the body of the \TT{try} statement is closed, $-1$ 
is written back to it.
We also see the addresses of filter and handler code.

Thus we can easily see the structure of the \IT{try/except} constructs in the function.\\
\\
Since the SEH setup code in the function prologue may be shared between many functions,
sometimes the compiler inserts a call to the \TT{SEH\_prolog()} function in the prologue, which does just that.

The SEH cleanup code is in the \TT{SEH\_epilog()} function.\\
\\
Let's try to run this example in \tracer{}:
\myindex{tracer}

\begin{lstlisting}
tracer.exe -l:2.exe --dump-seh
\end{lstlisting}

\begin{lstlisting}[caption=tracer.exe output]
EXCEPTION_ACCESS_VIOLATION at 2.exe!main+0x44 (0x401054) ExceptionInformation[0]=1
EAX=0x00000000 EBX=0x7efde000 ECX=0x0040cbc8 EDX=0x0008e3c8
ESI=0x00001db1 EDI=0x00000000 EBP=0x0018feac ESP=0x0018fe80
EIP=0x00401054
FLAGS=AF IF RF
* SEH frame at 0x18fe9c prev=0x18ff78 handler=0x401204 (2.exe!_except_handler3)
SEH3 frame. previous trylevel=0
scopetable entry[0]. previous try level=-1, filter=0x401070 (2.exe!main+0x60) handler=0x401088 (2.exe!main+0x78)
* SEH frame at 0x18ff78 prev=0x18ffc4 handler=0x401204 (2.exe!_except_handler3)
SEH3 frame. previous trylevel=0
scopetable entry[0]. previous try level=-1, filter=0x401531 (2.exe!mainCRTStartup+0x18d) handler=0x401545 (2.exe!mainCRTStartup+0x1a1)
* SEH frame at 0x18ffc4 prev=0x18ffe4 handler=0x771f71f5 (ntdll.dll!__except_handler4)
SEH4 frame. previous trylevel=0
SEH4 header:	GSCookieOffset=0xfffffffe GSCookieXOROffset=0x0
		EHCookieOffset=0xffffffcc EHCookieXOROffset=0x0
scopetable entry[0]. previous try level=-2, filter=0x771f74d0 (ntdll.dll!___safe_se_handler_table+0x20) handler=0x771f90eb (ntdll.dll!_TppTerminateProcess@4+0x43)
* SEH frame at 0x18ffe4 prev=0xffffffff handler=0x77247428 (ntdll.dll!_FinalExceptionHandler@16)
\end{lstlisting}

We see that the SEH chain consists of 4 handlers.\\
\\
\myindex{CRT}
The first two are located in our example. Two?
But we made only one?
Yes, another one has been set up in the \ac{CRT} function \TT{\_mainCRTStartup()}, and as it seems that it handles at least \ac{FPU} exceptions.
Its source code can found in the MSVC installation: \TT{crt/src/winxfltr.c}.\\
\\
The third is the SEH4 one in ntdll.dll, 
and the fourth handler is not MSVC-related and is located in ntdll.dll, and has a self-describing function name.\\
\\
As you can see, there are 3 types of handlers in one chain:

one is not related to MSVC at all (the last one) and two MSVC-related: SEH3 and SEH4.

\myparagraph{SEH3: two try/except blocks example}

\lstinputlisting[style=customc]{OS/SEH/2/3.c}

Now there are two \TT{try} blocks.
So the \IT{scope table} now has two entries, one for each block.
\IT{Previous try level} changes as execution flow enters or exits the \TT{try} block.

\lstinputlisting[caption=MSVC 2003,style=customasmx86]{OS/SEH/2/3_SEH3.asm}

If we set a breakpoint on the \printf{} function, which is called from the handler, 
we can also see how yet another SEH handler is added.

Perhaps it's another machinery inside the SEH handling process.
Here we also see our \IT{scope table} consisting of 2 entries.

\begin{lstlisting}
tracer.exe -l:3.exe bpx=3.exe!printf --dump-seh
\end{lstlisting}

\begin{lstlisting}[caption=tracer.exe output]
(0) 3.exe!printf
EAX=0x0000001b EBX=0x00000000 ECX=0x0040cc58 EDX=0x0008e3c8
ESI=0x00000000 EDI=0x00000000 EBP=0x0018f840 ESP=0x0018f838
EIP=0x004011b6
FLAGS=PF ZF IF
* SEH frame at 0x18f88c prev=0x18fe9c handler=0x771db4ad (ntdll.dll!ExecuteHandler2@20+0x3a)
* SEH frame at 0x18fe9c prev=0x18ff78 handler=0x4012e0 (3.exe!_except_handler3)
SEH3 frame. previous trylevel=1
scopetable entry[0]. previous try level=-1, filter=0x401120 (3.exe!main+0xb0) handler=0x40113b (3.exe!main+0xcb)
scopetable entry[1]. previous try level=0, filter=0x4010e8 (3.exe!main+0x78) handler=0x401100 (3.exe!main+0x90)
* SEH frame at 0x18ff78 prev=0x18ffc4 handler=0x4012e0 (3.exe!_except_handler3)
SEH3 frame. previous trylevel=0
scopetable entry[0]. previous try level=-1, filter=0x40160d (3.exe!mainCRTStartup+0x18d) handler=0x401621 (3.exe!mainCRTStartup+0x1a1)
* SEH frame at 0x18ffc4 prev=0x18ffe4 handler=0x771f71f5 (ntdll.dll!__except_handler4)
SEH4 frame. previous trylevel=0
SEH4 header:	GSCookieOffset=0xfffffffe GSCookieXOROffset=0x0
		EHCookieOffset=0xffffffcc EHCookieXOROffset=0x0
scopetable entry[0]. previous try level=-2, filter=0x771f74d0 (ntdll.dll!___safe_se_handler_table+0x20) handler=0x771f90eb (ntdll.dll!_TppTerminateProcess@4+0x43)
* SEH frame at 0x18ffe4 prev=0xffffffff handler=0x77247428 (ntdll.dll!_FinalExceptionHandler@16)
\end{lstlisting}

\myparagraph{SEH4}

\myindex{\BufferOverflow}
\myindex{Security cookie}
During a buffer overflow (\myref{subsec:bufferoverflow}) attack, the address of the \IT{scope table} 
can be rewritten, so starting from MSVC 2005, SEH3 was upgraded to SEH4 in order to have buffer overflow protection.
The pointer to the \IT{scope table} is now \glslink{xoring}{xored} with a \gls{security cookie}.
The \IT{scope table} was extended to have a header consisting of two pointers to \IT{security cookies}.

Each element has an offset inside the stack of another value: 
the address of the \gls{stack frame} (\EBP) \glslink{xoring}{xored} with the \TT{security\_cookie} , placed in the stack.

This value will be read during exception handling and checked for correctness.
The \IT{security cookie} in the stack is random each time, so hopefully a remote attacker can't predict it. \\
\\
The initial \IT{previous try level} is $-2$ in SEH4 instead of $-1$.

\def\SEHfour{1}
\input{OS/SEH/2/tikz}

Here are both examples compiled in MSVC 2012 with SEH4:

\lstinputlisting[caption=MSVC 2012: one try block example,style=customasmx86]{OS/SEH/2/2_SEH4.asm}

\lstinputlisting[caption=MSVC 2012: two try blocks example,style=customasmx86]{OS/SEH/2/3_SEH4.asm}

Here is the meaning of the \IT{cookies}: \TT{Cookie Offset} 
is the difference between the address of the saved EBP value in the stack
and the $EBP \oplus security\_cookie$ value in the stack.
\TT{Cookie XOR Offset} is an additional difference between the 
$EBP \oplus security\_cookie$ value and what is
stored in the stack.

If this equation is not true, the process is to halt due to stack corruption:

\begin{center}
$security\_cookie \oplus (CookieXOROffset + address\_of\_saved\_EBP) == stack[address\_of\_saved\_EBP + CookieOffset]$
\end{center}

If \TT{Cookie Offset} is $-2$, this implies that it is not present.

\myindex{tracer}
\IT{Cookies} checking is also implemented in my \tracer{},
see \href{http://go.yurichev.com/17061}{GitHub} for details.\\
\\
It is still possible to fall back to SEH3 in the compilers after 
(and including) MSVC 2005 by setting the \TT{/GS-} option,
however, the \ac{CRT} code use SEH4 anyway.

