\subsection{Simple example}

\lstinputlisting[style=customc]{patterns/04_scanf/1_simple/ex1.c}

It's not clever to use \scanf for user interactions nowadays. 
But we can, however, illustrate passing a pointer to a variable of type \Tint.

\subsubsection{About pointers}
\myindex{\CLanguageElements!\Pointers}

Pointers are one of the fundamental concepts in computer science.
Often, passing a large array, structure or object as an argument to another function is too expensive, while passing their address is much cheaper.
For example, if you going to print a text string to console, it's much easier to pass its address into \ac{OS} kernel.

In addition if the \gls{callee} function needs to modify something in the large array or structure received as a parameter and return back the entire structure then the situation is close to absurd.
So the simplest thing to do is to pass the address of the array or structure to the \gls{callee} function, and let it change what needs to be changed.

A pointer in \CCpp---is simply an address of some memory location.

\myindex{x86-64}
In x86, the address is represented as a 32-bit number (i.e., it occupies 4 bytes), while in x86-64 it is a 64-bit number (occupying 8 bytes).
By the way, that is the reason behind some people's indignation related to switching to x86-64---all pointers in the x64-architecture require twice as much space, including cache memory, which is ``expensive'' memory.

% TODO ... а делать разные версии memcpy для разных типов - абсурд
\myindex{\CStandardLibrary!memcpy()}
It is possible to work with untyped pointers only, given some effort; e.g. the standard C function \TT{memcpy()}, that copies a block from one memory location to another,
takes 2 pointers of type \TT{void*} as arguments, since it is impossible to predict the type of the data you would like to copy. Data types are not important, only the block size matters.

Pointers are also widely used when a function needs to return more than one value
(we are going to get back to this later
~(\myref{label_pointers})
).

\IT{scanf()} function---is such a case.

Besides the fact that the function needs to indicate how many values were successfully read, it also needs to return all these values.

In \CCpp the pointer type is only needed for compile-time type checking.

Internally, in the compiled code there is no information about pointer types at all.
% TODO это сильно затрудняет декомпиляцию

\input{patterns/04_scanf/1_simple/x86}
\input{patterns/04_scanf/1_simple/x64}
\input{patterns/04_scanf/1_simple/ARM}
\input{patterns/04_scanf/1_simple/MIPS/main}
