\section{System calls (syscall-s)}

\label{syscalls}
\myindex{syscall}

\myindex{kernel space}
\myindex{user space}
As we know, all running processes inside an \ac{OS} are divided into two categories:
those having full access to the hardware (\q{kernel space}) 
and those that do not (\q{user space}).

The \ac{OS} kernel and usually the drivers are in the first category.

All applications are usually in the second category.

\myindex{Glibc}
For example, Linux kernel is in \IT{kernel space}, but Glibc in \IT{user space}.

This separation is crucial for the safety of the \ac{OS}: it is very important not to give to any process the possibility to screw up
something in other processes or even in the \ac{OS} kernel.
\myindex{kernel panic}
\myindex{BSoD}
On the other hand, a failing driver or error inside the \ac{OS}'s kernel usually leads to a kernel panic or \ac{BSOD}.

The protection in the x86 processors allows to separate everything into 4 levels of protection (rings), but both in Linux
and in Windows only two are used: ring0 (\q{kernel space}) and ring3 (\q{user space}).

System calls (syscall-s)
are a point where these two areas are connected.

It can be said that this is the main \ac{API} provided to applications.

As in \gls{Windows NT}, the syscalls table resides in the \ac{SSDT}.

\myindex{Shellcode}

The usage of syscalls is very popular among shellcode and computer viruses authors, 
because it is hard to determine the addresses of
needed functions in the system libraries, but it is easier to use syscalls. However, much more code has to be
written due to the lower level of abstraction of the \ac{API}.

It is also worth noting that the syscall numbers may be different in various OS versions.

\subsection{Linux}
\label{linux_syscall}

\myindex{x86!\Instructions!INT!INT 0x80}
In Linux, a syscall is usually called via \TT{int 0x80}.
The call's number is passed in the \EAX register, and any other parameters~---in the other registers.

\lstinputlisting[caption=A simple example of the usage of two syscalls,style=customasmx86]{OS/linux_syscall.s}

Compilation:

\begin{lstlisting}
nasm -f elf32 1.s
ld 1.o
\end{lstlisting}

The full list of syscalls in Linux: \url{http://go.yurichev.com/17319}.

For system calls interception and tracing in Linux, strace(\myref{strace}) can be used.

\subsection{Windows}

\myindex{x86!\Instructions!INT!INT 0x2e}
\myindex{x86!\Instructions!SYSENTER}

Here they are called via \TT{int 0x2e} 
or using the special x86 instruction \TT{SYSENTER}.

The full list of syscalls in Windows: \url{http://go.yurichev.com/17320}.

Further reading:

\q{Windows Syscall Shellcode} by Piotr Bania: \url{http://go.yurichev.com/17321}.

