\mysection{64 bits}

\subsection{x86-64}
\myindex{x86-64}
\label{x86-64}

It is a 64-bit extension to the x86 architecture.

From the reverse engineer's perspective, the most important changes are:

\myindex{\CLanguageElements!\Pointers}
\begin{itemize}

\item

Almost all registers (except FPU and SIMD) were extended to 64 bits and got a R- prefix.
8 additional registers wer added.
Now \ac{GPR}'s are: \RAX, \RBX, \RCX, \RDX, 
\RBP, \RSP, \RSI, \RDI, \Reg{8}, \Reg{9}, \Reg{10}, 
\Reg{11}, \Reg{12}, \Reg{13}, \Reg{14}, \Reg{15}. 

It is still possible to access the \IT{older} register parts as usual. 
For example, it is possible to access the lower 32-bit part of the \RAX register using \EAX:

\RegTableOne{RAX}{EAX}{AX}{AH}{AL}

The new \GTT{R8-R15} registers also have their \IT{lower parts}: \GTT{R8D-R15D} (lower 32-bit parts),
\GTT{R8W-R15W} (lower 16-bit parts), \GTT{R8L-R15L} (lower 8-bit parts).

\RegTableFour{R8}{R8D}{R8W}{R8L}

The number of SIMD registers was doubled from 8 to 16: \XMM{0}-\XMM{15}.

\item

In Win64, the function calling convention is slightly different, somewhat resembling fastcall
(\myref{fastcall}).
The first 4 arguments are stored in the \RCX, \RDX, \Reg{8}, \Reg{9} registers, the rest~---in the stack.
The \gls{caller} function must also allocate 32 bytes so the \gls{callee} may save there 4 first arguments and use these 
registers for its own needs.
Short functions may use arguments just from registers, but larger ones may save their values on the stack.

System V AMD64 ABI (Linux, *BSD, \MacOSX)\SysVABI also somewhat resembles
fastcall, it uses 6 registers 
\RDI, \RSI, \RDX, \RCX, \Reg{8}, \Reg{9} for the first 6 arguments.
All the rest are passed via the stack.

See also the section on calling conventions~(\myref{sec:callingconventions}).

\item
The \CCpp \Tint type is still 32-bit for compatibility.

\item
All pointers are 64-bit now.

\end{itemize}

\myindex{Register allocation}

Since now the number of registers is doubled, the compilers have more space for maneuvering called 
\glslink{register allocator}{register allocation}.
For us this implies that the emitted code containing less number of local variables.

\myindex{DES}

For example, the function that calculates the first S-box of the DES encryption algorithm processes
32/64/128/256 values at once (depending on \GTT{DES\_type} type (uint32, uint64, SSE2 or AVX)) 
using the bitslice DES method
(read more about this technique here ~(\myref{bitslicedes})):

\lstinputlisting[style=customc]{patterns/20_x64/19_1.c}

There are a lot of local variables. 
Of course, not all those going into the local stack.
Let's compile it with MSVC 2008 with \GTT{/Ox} option:

\lstinputlisting[caption=\Optimizing MSVC 2008,style=customasmx86]{patterns/20_x64/19_2_msvc_Ox.asm}

5 variables were allocated in the local stack by the compiler.

Now let's try the same thing in the 64-bit version of MSVC 2008:

\lstinputlisting[caption=\Optimizing MSVC 2008,style=customasmx86]{patterns/20_x64/19_3_msvc_x64.asm}

Nothing was allocated in the local stack by the compiler, \GTT{x36} is synonym for \GTT{a5}.

\iffalse
% FIXME1 невнятно

By the way, we can see here that the function saved the \RCX and \RDX registers in space allocated by the \gls{caller},
but \Reg{8} and \Reg{9} were not saved but used from the beginning.
\fi

By the way, there are CPUs with much more \ac{GPR}'s, e.g. Itanium (128 registers).

\subsection{ARM}

64-bit instructions appeared in ARMv8.

\subsection{Float point numbers}

How floating point numbers are processed in x86-64 is explained here: \myref{floating_SIMD}.

\subsection{64-bit architecture criticism}

Some people has irritation sometimes: now one needs twice as much memory for storing pointers,
including cache memory, despite the fact that x64 \ac{CPU}s can address only 48 bits of external 
\ac{RAM}.

\begin{framed}
\begin{quotation}
Pointers have gone out of favor to the point now where I had to
flame about it because on my 64-bit computer that I have here, if I really
care about using the capability of my machine I find that I’d better not use
pointers because I have a machine that has 64-bit registers but it only has 2
gigabytes of RAM. So a pointer never has more than 32 significant bits to it.
But every time I use a pointer it’s costing me 64 bits and that doubles the
size of my data structure. Worse, it goes into the cache and half of my
cache is gone and that costs cash—cache is expensive.

So if I’m really trying to push the envelope now, I have to use arrays instead
of pointers. I make complicated macros so that it looks like I’m using
pointers, but I’m not really.
\end{quotation}
\end{framed}

( Donald Knuth in ``Coders at Work: Reflections on the Craft of Programming ''. )

Some people make their own memory allocators.
\myindex{CryptoMiniSat}
It's interesting to know about CryptoMiniSat\footnote{\url{https://github.com/msoos/cryptominisat/}} case.
This program rarely uses more than 4GiB of \ac{RAM}, but it uses pointers heavily.
So it requires less memory on 32-bit architecture than on 64-bit one.
To mitigate this problem, author made his own allocator (in \IT{clauseallocator.(h|cpp)} files),
which allows to have access to allocated memory using 32-bit identifiers instead of 64-bit pointers.

