\item[Jcc] (\RU{где}\EN{where} cc\EMDASH{}condition code)

\myindex{x86!\Instructions!JAE}
\myindex{x86!\Instructions!JA}
\myindex{x86!\Instructions!JBE}
\myindex{x86!\Instructions!JB}
\myindex{x86!\Instructions!JC}
\myindex{x86!\Instructions!JE}
\myindex{x86!\Instructions!JGE}
\myindex{x86!\Instructions!JG}
\myindex{x86!\Instructions!JLE}
\myindex{x86!\Instructions!JL}
\myindex{x86!\Instructions!JNAE}
\myindex{x86!\Instructions!JNA}
\myindex{x86!\Instructions!JNBE}
\myindex{x86!\Instructions!JNB}
\myindex{x86!\Instructions!JNC}
\myindex{x86!\Instructions!JNE}
\myindex{x86!\Instructions!JNGE}
\myindex{x86!\Instructions!JNG}
\myindex{x86!\Instructions!JNLE}
\myindex{x86!\Instructions!JNL}
\myindex{x86!\Instructions!JNO}
\myindex{x86!\Instructions!JNS}
\myindex{x86!\Instructions!JNZ}
\myindex{x86!\Instructions!JO}
\myindex{x86!\Instructions!JPO}
\myindex{x86!\Instructions!JP}
\myindex{x86!\Instructions!JS}
\myindex{x86!\Instructions!JZ}

\RU{Немало этих инструкций имеют синонимы (отмечены с AKA), это сделано для удобства}
\EN{A lot of these instructions have synonyms (denoted with AKA), this was done for convenience}.
\RU{Синонимичные инструкции транслируются в один и тот же опкод}
\EN{Synonymous instructions are translated into the same opcode}.
\RU{Опкод имеет т.н.}\EN{The opcode has a} \gls{jump offset}.

\label{Jcc}
\begin{description}
\item[JAE] \ac{AKA} JNC: \RU{переход если больше или равно (беззнаковый)}\EN{jump if above or equal (unsigned)}: CF=0
\EN{\item[JA] \ac{AKA} JNBE: jump if greater (unsigned): CF=0 and ZF=0}
\RU{\item[JA] \ac{AKA} JNBE: переход если больше (беззнаковый): CF=0 и ZF=0}
\EN{\item[JBE] jump if lesser or equal (unsigned): CF=1 or ZF=1}
\RU{\item[JBE] переход если меньше или равно (беззнаковый): CF=1 или ZF=1}
\item[JB] \ac{AKA} JC: \RU{переход если меньше (беззнаковый)}\EN{jump if below (unsigned)}: CF=1
\item[JC] \ac{AKA} JB: \RU{переход если CF=1}\EN{jump if CF=1}
\item[JE] \ac{AKA} JZ: \RU{переход если равно или ноль}\EN{jump if equal or zero}: ZF=1
\item[JGE] \RU{переход если больше или равно (знаковый)}\EN{jump if greater or equal (signed)}: SF=OF
\EN{\item[JG] jump if greater (signed): ZF=0 and SF=OF}
\RU{\item[JG] переход если больше (знаковый): ZF=0 и SF=OF}
\EN{\item[JLE] jump if lesser or equal (signed): ZF=1 or SF$\neq$OF}
\RU{\item[JLE] переход если меньше или равно (знаковый): ZF=1 или SF$\neq$OF}
\item[JL] \RU{переход если меньше (знаковый)}\EN{jump if lesser (signed)}: SF$\neq$OF
\item[JNAE] \ac{AKA} JC: \RU{переход если не больше или равно (беззнаковый)}\EN{jump if not above or equal (unsigned)} CF=1
\item[JNA] \RU{переход если не больше (беззнаковый)}\EN{jump if not above (unsigned)} CF=1 \AndENRU ZF=1
\item[JNBE] \RU{переход если не меньше или равно (беззнаковый)}\EN{jump if not below or equal (unsigned)}: CF=0 \AndENRU ZF=0
\item[JNB] \ac{AKA} JNC: \RU{переход если не меньше (беззнаковый)}\EN{jump if not below (unsigned)}: CF=0
\item[JNC] \ac{AKA} JAE: \RU{переход если CF=0, синонимично}\EN{jump CF=0 synonymous to} JNB.
\item[JNE] \ac{AKA} JNZ: \RU{переход если не равно или не ноль}\EN{jump if not equal or not zero}: ZF=0
\item[JNGE] \RU{переход если не больше или равно (знаковый)}\EN{jump if not greater or equal (signed)}: SF$\neq$OF
\EN{\item[JNG] jump if not greater (signed): ZF=1 or SF$\neq$OF}
\RU{\item[JNG] переход если не больше (знаковый): ZF=1 или SF$\neq$OF}
\item[JNLE] \RU{переход если не меньше (знаковый)}\EN{jump if not lesser (signed)}: ZF=0 \AndENRU SF=OF
\item[JNL] \RU{переход если не меньше (знаковый)}\EN{jump if not lesser (signed)}: SF=OF
\item[JNO] \RU{переход если не переполнение}\EN{jump if not overflow}: OF=0
\item[JNS] \RU{переход если флаг SF сброшен}\EN{jump if SF flag is cleared}
\item[JNZ] \ac{AKA} JNE: \RU{переход если не равно или не ноль}\EN{jump if not equal or not zero}: ZF=0
\item[JO] \RU{переход если переполнение}\EN{jump if overflow}: OF=1
\item[JPO] \RU{переход если сброшен флаг PF}\EN{jump if PF flag is cleared} (Jump Parity Odd)
\item[JP] \ac{AKA} \ac{JPE}: \RU{переход если выставлен флаг PF}\EN{jump if PF flag is set}
\item[JS] \RU{переход если выставлен флаг SF}\EN{jump if SF flag is set}
\item[JZ] \ac{AKA} JE: \RU{переход если равно или ноль}\EN{jump if equal or zero}: ZF=1
\end{description}

