\chapter{\PrintfSeveralArgumentsSectionName}

Now let's extend the \IT{\HelloWorldSectionName}~(\myref{sec:helloworld}) example, replacing \printf in
the \main function body with this:

\lstinputlisting[label=hw_c]{patterns/03_printf/1.c}

% sections
\input{patterns/03_printf/x86/main}
\ifdefined\IncludeARM
\input{patterns/03_printf/ARM/main}
\fi
\ifdefined\IncludeMIPS
\input{patterns/03_printf/MIPS/main}
\fi

\section{\Conclusion{}}

Here is a rough skeleton of the function call:

\lstinputlisting[caption=x86]{patterns/03_printf/skel1.lst.\LANG}

\lstinputlisting[caption=x64 (MSVC)]{patterns/03_printf/skel2.lst.\LANG}

\ifdefined\IncludeGCC
\lstinputlisting[caption=x64 (GCC)]{patterns/03_printf/skel3.lst.\LANG}
\fi

\ifdefined\IncludeARM
\lstinputlisting[caption=ARM]{patterns/03_printf/skel4.lst.\LANG}

\lstinputlisting[caption=ARM64]{patterns/03_printf/skel5.lst.\LANG}
\fi

\ifdefined\IncludeMIPS
\index{MIPS!O32}
\lstinputlisting[caption=MIPS (O32 calling convention)]{patterns/03_printf/skel_MIPS.lst.\LANG}
\fi

\section{By the way}

\index{fastcall}
By the way, this difference between the arguments passing in x86, x64, 
fastcall, ARM and MIPS is a good illustration of the fact that the CPU is oblivious to how the arguments are passed to functions. 
It is also possible to create a hypothetical compiler able to pass arguments 
via a special structure without using stack at all.

\ifdefined\IncludeMIPS
\index{MIPS!O32}
MIPS \$A0 \dots \$A3 registers are labelled this way only for convenience (that is in the O32 calling convention).
Programmers may use any other register (well, maybe except \$ZERO) 
to pass data or use any other calling convention.
\fi

The \ac{CPU} is not aware of calling conventions whatsoever.

We may also recall how newcoming assembly language programmers passing arguments into
other functions:
usually via registers, without any explicit order, or even via global variables.
Of course, it works fine.

