\subsection{MSVC: x86}

\lstinputlisting{patterns/04_scanf/2_global/ex2_MSVC.asm}

In this case the \TT{x} variable is defined in the \TT{\_DATA} segment and no memory is allocated in the local stack. It is accessed directly, not through the stack. 
Uninitialized global variables take no space in the executable file
(indeed, why one needs to allocate space for variables initially set to zero?), 
but when someone accesses their address, 
the \ac{OS} will allocate a block of zeroes there\footnote{That is how a \ac{VM} behaves}.

Now let's explicitly assign a value to the variable:

\lstinputlisting{patterns/04_scanf/2_global/default_value_EN.c}

We got:

\begin{lstlisting}
_DATA	SEGMENT
_x	DD	0aH

...
\end{lstlisting}

Here we see a value \TT{0xA} of DWORD type (DD stands for DWORD = 32 bit) for this variable.

If you open the compiled .exe in \IDA, you can see the \IT{x} variable placed at the beginning of 
the \TT{\_DATA} segment, and after it you can see text strings.

If you open the compiled .exe from the previous example in \IDA, where the value of \IT{x} was not set, you would see something like this:

\begin{lstlisting}
.data:0040FA80 _x              dd ?                    ; DATA XREF: _main+10
.data:0040FA80                                         ; _main+22
.data:0040FA84 dword_40FA84    dd ?                    ; DATA XREF: _memset+1E
.data:0040FA84                                         ; unknown_libname_1+28
.data:0040FA88 dword_40FA88    dd ?                    ; DATA XREF: ___sbh_find_block+5
.data:0040FA88                                         ; ___sbh_free_block+2BC
.data:0040FA8C ; LPVOID lpMem
.data:0040FA8C lpMem           dd ?                    ; DATA XREF: ___sbh_find_block+B
.data:0040FA8C                                         ; ___sbh_free_block+2CA
.data:0040FA90 dword_40FA90    dd ?                    ; DATA XREF: _V6_HeapAlloc+13
.data:0040FA90                                         ; __calloc_impl+72
.data:0040FA94 dword_40FA94    dd ?                    ; DATA XREF: ___sbh_free_block+2FE
\end{lstlisting}

\TT{\_x} is marked with \TT{?} with the rest of the variables that do not need to be initialized. 
This implies that after loading the .exe to the memory, a space for all these variables is to be 
allocated and filled with zeroes \cite[6.7.8p10]{C99TC3}. 
But in the .exe file these uninitialized variables do not occupy anything.
This is convenient for large arrays, for example.

\ifdefined\IncludeOlly
\input{patterns/04_scanf/2_global/olly.tex}
\fi

\ifdefined\IncludeGCC
\subsection{GCC: x86}

\myindex{ELF}
The picture in Linux is near the same, with the difference that the uninitialized variables are located in the \TT{\_bss} segment. 
In \ac{ELF} file this segment has the following attributes:

\begin{lstlisting}
; Segment type: Uninitialized
; Segment permissions: Read/Write
\end{lstlisting}

If you, however, initialise the variable with some value e.g. 10, 
it is to be placed in the \TT{\_data} segment, which has the following attributes:

\begin{lstlisting}
; Segment type: Pure data
; Segment permissions: Read/Write
\end{lstlisting}
\fi

\subsection{MSVC: x64}

\lstinputlisting[caption=MSVC 2012 x64]{patterns/04_scanf/2_global/ex2_MSVC_x64_EN.asm}

The code is almost the same as in x86.
Please note that the address of the $x$ variable is passed to \TT{scanf()} using a \LEA instruction,
while the variable's value is passed to the second \printf using a \MOV instruction.
\TT{DWORD PTR}---is a part of the assembly language (no relation to the machine code),
indicating that the variable data size is 32-bit and the \MOV instruction has to be encoded accordingly.

