\ifdefined\ENGLISH
\newcommand{\GlobalVarsSectionName}{Global variables}
\sectionold{\GlobalVarsSectionName}
\myindex{\GlobalVarsSectionName}
\label{scanf_global_variable}

What if the \TT{x} variable from the previous example was not local but a global one? 
Then it would have been accessible from any point, not only from the function body. 
Global variables are considered \gls{anti-pattern}, but for the sake of the experiment, we could do this.

\lstinputlisting{patterns/04_scanf/2_global/ex2_EN.c}
\fi

\ifdefined\ITALIAN
\newcommand{\GlobalVarsSectionName}{Varibili globali}
\sectionold{\GlobalVarsSectionName}
\myindex{\GlobalVarsSectionName}
\label{scanf_global_variable}

E se la variabile \TT{x} dell'esempio precedente non fosse locale ma globale? 
Sarebbe stata accessibile da qualunque punto del codice, non soltanto nel corpo della funzione. 
Le variabili glocali sono considerate \gls{anti-pattern}, ma per il nostro interesse di sperimentare ci è concesso usarle.

\lstinputlisting{patterns/04_scanf/2_global/ex2_EN.c}
\fi

\ifdefined\RUSSIAN
\newcommand{\GlobalVarsSectionName}{Глобальные переменные}
\sectionold{\GlobalVarsSectionName}
\myindex{\GlobalVarsSectionName}
\label{scanf_global_variable}

А что если переменная \TT{x} из предыдущего примера будет глобальной переменной, а не локальной? 
Тогда к ней смогут обращаться из любого другого места, а не только из тела функции. 
Глобальные переменные считаются \glslink{anti-pattern}{анти-паттерном},
но ради примера мы можем себе это позволить.

\lstinputlisting{patterns/04_scanf/2_global/ex2_RU.c}
\fi

\ifdefined\BRAZILIAN
\newcommand{\GlobalVarsSectionName}{Variáveis globais}
\sectionold{\GlobalVarsSectionName}
\myindex{\GlobalVarsSectionName}
\label{scanf_global_variable}

E se a variável \TT{x} do último exemplo não fosse local, mas sim global?
Então ela teria que ser acessível de qualquer ponto, não somente pelo corpo da função.
Variáveis globais são consideradas maus hábitos, mas pelo bem do experimento, nós faremos isso.

% TODO translate
\lstinputlisting{patterns/04_scanf/2_global/ex2_EN.c}
\fi


\EN{\input{patterns/04_scanf/2_global/ex2_global_vars_x86_EN}}
\RU{\input{patterns/04_scanf/2_global/ex2_global_vars_x86_RU}}
\PTBR{\input{patterns/04_scanf/2_global/ex2_global_vars_x86_PTBR}}
\EN{\input{patterns/04_scanf/2_global/ex2_global_vars_ARM_EN}}
\RU{\input{patterns/04_scanf/2_global/ex2_global_vars_ARM_RU}}
\input{patterns/04_scanf/2_global/MIPS/main}
