\subsubsection{MSVC}

\RU{Компилируем в}\EN{Compile it in} MSVC 2010:

\lstinputlisting[caption=MSVC 2010: \ttf{}]{patterns/12_FPU/1_simple/MSVC.asm.\LANG}

\RU{\FLD берет 8 байт из стека и загружает их в регистр \ST{0}, автоматически конвертируя во внутренний 
80-битный формат (\IT{extended precision}).}
\EN{\FLD takes 8 bytes from stack and loads the number into the \ST{0} register, automatically converting 
it into the internal 80-bit format (\IT{extended precision}).}

\index{x86!\Instructions!FDIV}
\RU{\FDIV делит содержимое регистра \ST{0} на число, лежащее по адресу \TT{\_\_real@40091eb851eb851f}~--- 
там закодировано значение 3,14. Синтаксис ассемблера не поддерживает подобные числа, 
поэтому мы там видим шестнадцатеричное представление числа 3,14 в формате IEEE 754.}
\EN{\FDIV divides the value in \ST{0} by the number stored at address 
\TT{\_\_real@40091eb851eb851f}~---the value 3.14 is encoded there. 
The assembly syntax doesn't support floating point numbers, so 
what we see here is the hexadecimal representation of 3.14 in 64-bit IEEE 754 format.}

\RU{После выполнения \FDIV в \ST{0} остается \glslink{quotient}{частное}.}
\EN{After the execution of \FDIV \ST{0} holds the \gls{quotient}.}

\index{x86!\Instructions!FDIVP}
\RU{Кстати, есть ещё инструкция \FDIVP, которая делит \ST{1} на \ST{0}, 
выталкивает эти числа из стека и заталкивает результат. 
Если вы знаете язык Forth\FNURLFORTH, то это как раз оно и есть~--- стековая машина\FNURLSTACK.}
\EN{By the way, there is also the \FDIVP instruction, which divides \ST{1} by \ST{0}, 
popping both these values from stack and then pushing the result. 
If you know the Forth language\FNURLFORTH,
you can quickly understand that this is a stack machine\FNURLSTACK.}

\RU{Следующая \FLD заталкивает в стек значение $b$.}
\EN{The subsequent \FLD instruction pushes the value of $b$ into the stack.}

\RU{После этого в \ST{1} перемещается результат деления, а в \ST{0} теперь $b$.}
\EN{After that, the quotient is placed in \ST{1}, and \ST{0} has the value of $b$.}

\index{x86!\Instructions!FMUL}
\RU{Следующий \FMUL умножает $b$ из \ST{0} на значение \TT{\_\_real@4010666666666666}~--- 
там лежит число 4,1~--- и оставляет результат в \ST{0}.}
\EN{The next \FMUL instruction does multiplication: $b$ from \ST{0} is multiplied by by value at 
\TT{\_\_real@4010666666666666} (the numer 4.1 is there) and leaves the result in the \ST{0} register.}

\index{x86!\Instructions!FADDP}
\RU{Самая последняя инструкция \FADDP складывает два значения из вершины стека 
в \ST{1} и затем выталкивает значение, лежащее в \ST{0}. 
Таким образом результат сложения остается на вершине стека в \ST{0}.}
\EN{The last \FADDP instruction adds the two values at top of stack, storing the result in \ST{1} 
and then popping the value of \ST{0}, thereby leaving the result at the top of the stack, in \ST{0}.}

\RU{Функция должна вернуть результат в \ST{0}, так что больше ничего здесь не производится, 
кроме эпилога функции.}
\EN{The function must return its result in the \ST{0} register, 
so there are no any other instructions except the function epilogue after \FADDP.}

\ifdefined\IncludeOlly
\input{patterns/12_FPU/1_simple/olly.tex}
\fi
